Lattice Boltzmann Microchannel Simulations via Binary Liquid Model

### 2011-04-06 Conference/CFD2011/paper.tex
Started writing it as a paper. Need to do the introduction.

\title{Multiphase microchannel flows simulations via the lattice Boltzmann scheme}
\paperID{CFD11-52}
\author{Alexandr}{Kuzmin} %{forename}{surname}
\presenting  %the previous author is presenting the paper (name becomes underlined)
\email{shurik.kuzmin@gmail.com}%e-mail address of the previous author
\address{Chemical and Materials Engineering, University of Alberta, 9107-116 Street,  T6G 2V4
Edmonton, CANADA}%affiliation of the previous author
\author{Dmitry}{Eskin}
\email{deskin@slb.com}
\address{DBR Schlumberger Technology Center, 9450 17 Ave NW, T6N 1M9 Edmonton,CANADA}
\author{Jos}{Derksen}
%\address{NTNU Department of Physics, 7491 Trondheim, CANADA}
\email{jos@ualberta.ca}
\address{Chemical and Materials Engineering, University of Alberta, 9107-116 Street,  T6G 2V4
Edmonton, CANADA}%affiliation of the previous author

@@

\abstract{
In this work a free-energy binary liquid lattice Boltzmann scheme is used to simulate
Taylor/Bretherton flow in a micro-channel where elongated gas bubbles move through a liquid with
thin liquid films between the bubbles and the channel walls. The main focus of our work is a
feasibility study of the binary-liquid model lattice Botlzmann model to simulate the
Bretherton/Taylor phenomena. The numerical scheme has a diffuse
interface, and we thoroughly assess resolution requirements for correctly resolving
the liquid film and bubble motion. The simulations are two-dimensional and three-dimensional and
span a capillary number
range of $0.03$ to $0.8$ (for 2D) and range of $0.1$ to $1.0$ where the capillary number is based
on the liquid's viscosity, the
velocity of the bubble's interface, and the interfacial tension. The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. We obtain grid independent
results
as long as the liquid film thickness is at least two times the width of the diffuse interface. The
results show that the lattice Boltzmann free energy binary liquid model can be used to simulate the
Bretherton problem with a good accuracy.
}
\keywords{CFD, Multiphase Flows, Lattice Boltzmann Method, Microchannel simulations, Bretherton
Problem, Binary liquid, Free energy method}
\normalfont\normalsize

@@

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
narrow channels. Depending on the geometry it was found that the deposited film thickness
is a complicated function of the capillary number $Ca$, see Eq. (\ref{capillary:number:definition}).
For example, the deposited film thickness
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers for circular channels
\cite{bretherton,heil-bretherton}. 
\vskip .4em
The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
is of significant interests for chemical industry as far as it is widely used in chemical monolith
microreactors \cite{kreutzer-pressure-drop}. While it
is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, it's
not possible to extrapolate it to the wider range of capillary numbers used in the chemical
industry. Thus, the problem of consistent numerical simulations arises.
\vskip .4em
For what follows we distinguish two- and three-dimensional cases and give a detailed description
of the literature review. 
\begin{description}
\item[2D case]
Bretherton made analytical assymptotic analysis and predicted the $\Ca^{2/3}$ dependency of the film
thickness for the circular shaped microchannels in the range of the small capillary numbers
(typically $\Ca<0.003$), where capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
\Ca=\frac{\mu_{\mathrm{liq}} U_\mathrm{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is the bubble velocity, and
$\gamma$ is the interfacial tension between gas and liquid. For the case of the continuous
interface models,  $U_{bubble}$ is the interface velocity.

@@

To consistently  predict a flow pattern for capillaries in
different ranges of the capillary number, simulations have been conducted. The simulations are
typically validated with analytical solutions in the range of the small capillary numbers
 and then extende to more practical regimes.

@@

The experimental works supported by numerical simulations reveal interesting phenomena happening in
three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for the
certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case,
Fig. \ref{fig:sym:antisym}.
From herein, we limit ourselves to the case of microchannels with square crosssections.
Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like as the rounded square. The dependance
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the flow becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the film
thickness dependency on the capillary number is in Fig. \ref{fig:heil:three:dim}.

\citet{heil-threedim} also indicated another interesting phenomena. They indicated the transition
capillary number where the streamlines pattern
changes from having vortex in front of the bubble and not having it for larger capilllary number,
Fig. \ref{fig:streamlines:pattern}.
For the square channel the critical number is $Ca=0.691$.   
%\begin{figure}
%\includegraphics[width=0.5\textwidth]{Figures/ca05_clean.eps}\\
%\includegraphics[width=0.5\textwidth]{Figures/ca10_clean.eps}
%\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
%Streamlines in the moving frame of reference for a tube with $Ca = 0.5$ (top) and $Ca = 2.0$
%(bottom). 
%\label{fig:heil:recirculating}}
%\end{figure}

@@

circular shaped capillaries. They also found the empirical correlation which allows to put

@@

interface method based on the non-uniform grid. It is of particular interest for us to compare it

@@

numerical simulations for the determination of the critical capillary number $\hat{Ca}$.

@@

are performed with the help of the boundary value approach

@@

droplet breakup. The continuous interface models are more flexible for those kind of simulations.

@@

{\emph goal} of this work is to reproduce the numerical results and phenomena presented above
with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We do
a particular focus on exploring the parameter range of the binary liquid lattice
Boltzmann method for correctly resolving the flow in a range of capillary numbers. A thorough
parametric study for this approach and the ability of the
method to reconstruct consistent with other methods results has not been done to the best of the
authors' knowledge.

@@

Therefore, the major governing factor for microchannel flows is not the density ratio, but the
viscosity ratio. The goal of this work is to do a feasibility assessment study of the LBM
binary-liquid model to simulate and correctly predict flow patterns for th Bretherton/Taylor

@@

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

@@

Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow between plates and three-dimensional square
shaped microchannels, both driven by a body force. We limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability. This
implies that we can use the periodic boundary conditions in the streamwise direction. As soon as the
periodic boundary
conditions are applied, not a single bubble but a bubble train is simulated. In this case one
needs
to ensure that the distance between bubbles is large enough to exclude mutual bubble influence. 
Therefore, we chose the bubble length to be $5$ channel heights. This number is shown
to be sufficient to give results consistent with the theory. The
channel length is taken to be $3$ bubble lengths (or $15$ channel heights)
to minimize influence of one bubble on another,
because of the periodicity of boundary conditions. The geometry
dimensions are represented in Fig. \ref{fig:benchmark:sketch}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/benchmark.eps}
\caption{The two-dimensional benchmark sketch. $\delta$ corresponds to film thickness. $5\xi$
corresponds to the interface thickness. We take the microchannel length to be $15$ times larger than
its height. The threedimensional benchmark has the same geometry extending one dimension to have a
square shaped crosssection.
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{2D results}
Two-dimensional binary-liquid simulations are presented in our recent work
\cite{kuzmin-binary2d}. Here we outline two most important results. The first one addresses the
question of how the interface should be resolved to obtain grid independent result. Though the
simulations are performed for the lattice Boltzmann model, the results can be applied for any
uniform grid continuous interface method. The second part of results are comparisons with the
literature results for the film thickness dependency on the capillary number.

@@

We performed a number of simulations for grids indicated in
Table \ref{table:parameters:grid:refinement}. Other parameters are defined as $k=0.04$, $A=0.04$ and
$\Gamma=1.0$. Note that in the case of the half-way bounce-back walls \cite{yu} which are used in
the
simulations one needs to calculate the film thickness as:
\begin{equation}
\delta=(\phi_0-0.5)/(N_y-2)=(\phi_0-0.5)/H_{\mathrm{eff}},
\end{equation}
where $\phi_0$ is the grid coordinate where phase field is $0$, $H_{\mathrm{eff}}=N_y-2$
is the effective channel height.
If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions between the grid nodes
which represent
the physical domain, top and bottom nodes are bounce-back nodes. Then the effective wall location is
in the middle between bounce-back and fluid
nodes. Overall there are $N_y-2$ nodes representing the fluid. Note that it is a big
simplification to impose the boundary in the middle between the bounce-back
node and the fluid node. The location of the wall for bounce-back nodes is
viscosity dependent \cite{ginzburg-multireflection}.  However, the effective location of the
wall for the
multiphase models to the best authors' knowledge is not yet derived.
The bubble is initialized as a rectangular box with coordinates
$y=7\times\frac{H_\mathrm{eff}}{100}\dots N_y-7\times\frac{H_\mathrm{eff}}{100}-1$,
$x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
$\phi_{\mathrm{bubble}}=-1$. All other nodes are initialized with the phase field
$\phi=1$. The initialization procedure is done to keep the self-similarity. The initial width and
body force were chosen to aim for the capillary number $\Ca=0.05$ \cite{kuzmin-binary2d}.
After choosing the reference parameters, the grid refinement procedure
needs to keep the macroscopic parameters constant.  It is easy to check
that it yields the following quantity to
be independent of grid size
$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = \mathrm{const}$. For the grid $H_{\mathrm{eff}}=100$ we chose the body force as
$\frac{\mathrm{d}P}{\mathrm{d}x}=1.508 \times 10^{-6}$ lattice units.
The
simulation results in terms of the film thickness $\delta$, interface bubble
velocity $U_{bubble}$ and associated capillary numbers are summarized in Table
\ref{table:parameters:grid:refinement}. 
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see that results converge for $H_{\mathrm{eff}}\geq 175$. To calculate how well the interface is
resolved, the ratio of the interface thickness to the film thickness is calculated. The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. The ratio of the interface thickness to the film thickness
$5\xi/H_{\mathrm{film}}$ is shown in Table \ref{table:parameters:grid:refinement}.

@@

\subsubsection{Capillary number region}
The purpose of this section is to validate the correlations of
\citet{giavedoni-numerical} and
\citet{heil-bretherton} for a film thickness dependancy on the range of capillary
numbers. Because of limited computational resources, we skip the
small capillary numbers and make calculations for the range of capillary numbers $0.03-0.8$, which
is a computationally reasonable task.  To be consistent with the grid independant results, we choose
the grid to be
$202 \times 3001$. Then $5$ lattice Boltzmann units do not occupy more than $60$
percent of the effective film thickness. Because the simulation gets unstable with
smaller grids and larger gradients, all the capillary number simulations were
performed on the same grid. To properly initialize the body force, the
proportionality law was utilized. The forcing
$6 \cdot 10^{-6}/16$ was chosen to obtain the predicted capillary
number $0.05$. The forcing for other grids can be obtained using
simple proportionality relationships:
\begin{equation}
\begin{aligned}
&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\
&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{N_y^2},
\end{aligned}
\end{equation}
where the subscript ,,lit'' stands for the predicted capillary number
\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
involved because all simulations are conducted on the same grid.
The pressure gradient can be obtained through the capillary number
ratio:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d} x}=6 \cdot 10^{-6}/16 \frac{\Ca_{\mathrm{lit}}}{0.05}
\end{equation}
The film thickness is initialized through the ratio of capillary numbers as well:
\begin{equation*}
w=12 \frac{\Ca_{\mathrm{lit}}}{0.05}
\end{equation*}
The results obtained after $2\cdot10^5$ steps are presented in Table
\ref{table:parameters:capillary:number}. One can see that the calculated
capillary numbers are overpredicted but the obtained thicknesses are overpredicted
as well due to the body force setting caused by the Poiseuille flow assumption. If one
needs to obtain the desired capillary number, the shooting method for the forcing is
necessary, starting with the Poiseuille pressure gradient force as the initial condition. We
extracted
the data of 
\citet{giavedoni-numerical} and \citet{heil-bretherton} with the help of software ``Engauge
Digitizer'' and compared them with our results (see 
Fig. \ref{fig:capillary:comparison}).
\begin{figure}
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison.eps}
\caption{A comparison between simulation results and results of
\citet{giavedoni-numerical} and \citet{heil-bretherton}. One can see a
reasonable agreement. The plots depict the film thickness as a function of the
the capillary number.\label{fig:capillary:comparison}}
\end{figure}

@@

\subsection{3D results}
One can see from Fig. \ref{fig:heil:three:dim} that the separation of the axial and diagonal radii
happens at the capillary numbers less than $0.1$. However the axial radius has a value around $0.99
H_{\mathrm{eff}}$. Therefore, the film occupies $0.01 H_{\mathrm{eff}}$ which needs to be resolved
with at minimum $10$ lattice units (see results for grid independency). This condition implies grids
to be of at least size $1000\times1000\times 15000$, which is not a computationally feasible
task. Even the symmetric and antisymmetric profiles are obtained by current simulations, Fig.
\ref{fig:sym:antisym}, the grid independence is not guaranteed.
Therefore our target simulations are from $\Ca\geq 0.1$, where the axial radius equals to the
diagonal radius. In what follows we will show two particular simulations. One is for the dependance
of the bubble radius on the capillary number. Another one is for the velocity patterns. In
comparison with simulations of \citet{heil-threedim} the transition capillary
number is located in the range $0.47<\Ca<0.63$. 
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/sym.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/antisym.eps}\\
\caption{Symmetric ($\Ca=0.63$) and antisymmetric ($\Ca=0.026$) crosssections in the middle of the
bubble. One can see that for the symmetric profile $R_h=R_d$ and for the antisymmetric $R_h\neq
R_d$.  \label{fig:sym:antisym} }
\end{figure}

We also simulate only the quarter of the channel by imposing the mirror boundary conditions. This
gives the improvement in memory size and time by factor of four.  
\subsubsection{Capillary region}
We performed a number of simulations with CPU code of grid size $52\times52\times 1500$ for the
quarter of geometry and of refined grid size $82\times82\times 1500$. All other parameters were
chosen as
$k=0.04$,$A=0.04$, and $\Gamma=1.0$.

>>>>

\title{Lattice Boltzmann Microchannel Simulations via Binary Liquid Model}
\paperID{CFD11-52}
\author{Alexandr}{Kuzmin} %{forename}{surname}
\presenting  %the previous author is presenting the paper (name becomes underlined)
\email{kuzmin@ualberta.ca}%e-mail address of the previous author
\address{Chemical and Materials Engineering, University of Alberta, 9107-116 Street,  T6G 2V4
Edmonton, CANADA}%affiliation of the previous author
\author{Dmitry}{Eskin}
\email{deskin@slb.com}
\address{DBR Schlumberger Technology Center, 9450 17 Ave NW, T6N 1M9 Edmonton,CANADA}
\author{Jos}{Derksen}
%\address{NTNU Department of Physics, 7491 Trondheim, CANADA}
\email{jos@ualberta.ca}
\address{Chemical and Materials Engineering, University of Alberta, 9107-116 Street,  T6G 2V4
Edmonton, CANADA}%affiliation of the previous author

@@

\abstract{
In this work a free-energy binary liquid lattice Boltzmann scheme is used to simulate
Taylor/Bretherton flow in a micro-channel where elongated gas bubbles move through a liquid with
thin liquid films between the bubbles and the channel walls. The main focus of our work is a
feasibility study of the binary-liquid model lattice Boltzmann model to simulate the
Bretherton/Taylor phenomena. The numerical scheme has a diffuse
interface, and we thoroughly assess resolution requirements for correctly resolving
the liquid film and bubble motion. The simulations are two-dimensional and three-dimensional and
span a capillary number
range of $0.03$ to $0.8$ (for 2D) and range of $0.1$ to $1.0$ where the capillary number is based
on the liquid's viscosity, the
velocity of the bubble's interface, and the interfacial tension. The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. We obtain grid independent
results
as long as the liquid film thickness is at least two times the width of the diffuse interface. The
results show that the lattice Boltzmann free energy binary liquid model can be used to simulate the
Bretherton problem with good accuracy.
}
\keywords{CFD, Multiphase Flows, Lattice Boltzmann Method, Microchannel flow, Bretherton
problem, Binary liquid, Free energy method}
\normalfont\normalsize

@@

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
narrow channels. Depending on the geometry it was found that the deposited film thickness
is a complicated function of the capillary number $Ca$, Eq. (\ref{capillary:number:definition}).
For example, the deposited film thickness
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers for circular channels
\cite{bretherton,heil-bretherton}. 
\vskip .4em
The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
is of significant interests for chemical industry as it is widely used in chemical monolith
microreactors \cite{kreutzer-pressure-drop}. While it
is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, it's
not possible to extrapolate it to the wider range of capillary numbers used in the chemical
industry. Thus, the desire of consistent numerical simulations arises.
\vskip .4em
For what follows we distinguish two- and three-dimensional cases and give a detailed acount
of the literature. 
\begin{description}
\item[2D case]
Bretherton made analytical assymptotic analysis and predicted the $\Ca^{2/3}$ dependency of the film
thickness for the circular shaped microchannels in the range of the small capillary numbers
(typically $\Ca<0.003$), where capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
\Ca=\frac{\mu_{\mathrm{liq}} U_\mathrm{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid's kinematic viscosity, $U_{bubble}$ is the bubble velocity, and
$\gamma$ is the interfacial tension between gas and liquid. For the case of the continuous
interface models,  $U_{bubble}$ is the interface velocity.

@@

To consistently  predict a flow pattern for capillaries in
different ranges of the capillary number, simulations have been conducted. The simulations are
typically validated with analytical solutions in the range of the small capillary numbers
 and then extended to more practical regimes.

@@

The experimental works supported by numerical simulations reveal interesting phenomena happening in
three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case,
Fig. \ref{fig:sym:antisym}.
From hereon, we limit ourselves to the case of microchannels with square crosssections.
Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

\citet{heil-threedim} also indicated another interesting phenomenon. They indicated the transition
capillary number where the streamlines pattern
changes from having a vortex in front of the bubble and not having it for larger capilllary
numbers,
Fig. \ref{fig:streamlines:pattern}.
For the square channel the critical number is $Ca=0.691$.   
%\begin{figure}
%\includegraphics[width=0.5\textwidth]{Figures/ca05_clean.eps}\\
%\includegraphics[width=0.5\textwidth]{Figures/ca10_clean.eps}
%\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
%Streamlines in the moving frame of reference for a tube with $Ca = 0.5$ (top) and $Ca = 2.0$
%(bottom). 
%\label{fig:heil:recirculating}}
%\end{figure}

@@

circular shaped capillaries. They also found an empirical correlation which allows to put

@@

interface method based on a non-uniform grid. It is of particular interest for us to compare it

@@

numerical simulations for the determination of the critical capillary number $\widehat{Ca}$.

@@

are performed with the help of the finite element method

@@

droplet breakup. The continuous interface models are more flexible for this kind of simulations.

@@

\emph{goal} of this work is to reproduce the numerical results and phenomena presented above
with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We
focus on exploring the parameter range of the binary liquid lattice
Boltzmann method for correctly resolving the flow in a range of capillary numbers. 

@@

Therefore, a major governing factor for microchannel flows is not the density ratio, but the
viscosity ratio. The goal of this work is to do a feasibility study of the LBM
binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor

@@

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

@@

Given all the concerns and challenges, the suggested lattice Boltzmann framework
is a two-dimensional flow between plates and three-dimensional square
shaped microchannels, both driven by a body force. We limit ourselves to the study
of body force driven flows, because of their simplicity and better numerical stability. 
This
implies that we can use the periodic boundary conditions in the streamwise direction. As soon as the
periodic boundary
conditions are applied, not a single bubble but a bubble train is simulated. In this case one
needs
to ensure that the channel length is large enough to exclude mutual bubble influence.  The
channel length is taken to be $15$ channel heights with the bubble length to be $5$ channel
heights. The chosen numbers are shown to be sufficient to give results consistent with the theory,
i.e. for the the film thickness to stabilize and to exclude mutual bubble influence. The geometry
dimensions are represented in Fig. \ref{fig:benchmark:sketch}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/benchmark.eps}
\caption{The $x-y$ benchmark crosssection. $\delta$ corresponds to the film thickness. $5\xi$
corresponds to the interface thickness. We take the microchannel length to be $15$ times larger than
its height. The threedimensional benchmark has the same geometry extending it in the $z$ dimension
to have a
square shaped crosssection.
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{2D results}
Two-dimensional binary-liquid simulations are presented in our recent work
\cite{kuzmin-binary2d}. Here we outline the two most important results. The first one addresses the
question of how the interface should be resolved to obtain grid independent result. Though the
simulations are performed for the lattice Boltzmann model, the results can be applied for any
uniform grid continuous interface method. The second part of results are comparisons with the
literature for the film thickness dependency on the capillary number.

@@

We performed a number of simulations for grids indicated in
Table \ref{table:parameters:grid:refinement}. Other parameters are defined as $k=0.04$, $A=0.04$ and
$\Gamma=1.0$. Note that in the case of the half-way bounce-back walls \cite{yu} which are used in
the
simulations one needs to calculate the film thickness as:
\begin{equation}
\delta=(\phi_0-0.5)/(N_y-2)=(\phi_0-0.5)/H_{\mathrm{eff}},
\end{equation}
where $\phi_0$ is the grid coordinate where phase field is $0$, $H_{\mathrm{eff}}=N_y-2$
is the effective channel height.
If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions representing
the physical domain with top and bottom nodes being bounce-back nodes. Then the effective wall
location is
in the middle between bounce-back and fluid
nodes. Overall there are $N_y-2$ nodes representing the fluid. Note that it is a simplification to
impose the boundary in the middle between the bounce-back
node and the fluid node. The location of the wall for bounce-back nodes is
viscosity dependent \cite{ginzburg-multireflection}.  However, the effective location of the
wall for the
multiphase models to the best authors' knowledge is not yet derived.
The bubble is initialized as a rectangular box with coordinates
$y=7\times\frac{H_\mathrm{eff}}{100}\dots N_y-7\times\frac{H_\mathrm{eff}}{100}-1$,
$x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
$\phi_{\mathrm{bubble}}=-1$. All other nodes are initialized with the phase field
$\phi=1$. The initialization procedure is done to keep the self-similarity. The initial width and
body force were chosen to aim for the capillary number $\Ca=0.05$ \cite{kuzmin-binary2d}.
After choosing the reference parameters, the grid refinement procedure
needs to keep the macroscopic parameters constant.  It is easy to check
that it yields the following quantity to
be independent of grid size
$H_{\mathrm{eff}}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = \mathrm{const}$. For the grid $H_{\mathrm{eff}}=100$ we chose the body force as
$\frac{\mathrm{d}P}{\mathrm{d}x}=1.508 \times 10^{-6}$ lattice units.
The
simulation results in terms of the film thickness $\delta$, interface bubble
velocity $U_{bubble}$ and associated capillary numbers are summarized in Table
\ref{table:parameters:grid:refinement}. 
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see that results converge for $H_{\mathrm{eff}}\geq 175$. To calculate how well the interface is
resolved, the ratio of the interface thickness to the film thickness is calculated. The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. The ratio of the interface thickness to the film thickness
$5\xi/H_{\mathrm{film}}$ is shown in Table \ref{table:parameters:grid:refinement}.

@@

\subsubsection{Capillary number region}
The purpose of this section is to validate the correlations of
\citet{giavedoni-numerical} and
\citet{heil-bretherton} for a film thickness dependancy on the range of capillary
numbers. Because of limited computational resources, we skip the
small capillary numbers and make calculations for the range of capillary numbers $0.03-0.8$, which
is a computationally reasonable task.  To be consistent with the grid independant results, we choose
the grid to be
$202 \times 3001$. Then $5$ lattice Boltzmann units do not occupy more than $60$
percent of the effective film thickness. Because the simulation gets unstable with
smaller grids and larger gradients, all the capillary number simulations were
performed on the same grid. To properly initialize the body force, the
proportionality law was utilized. The forcing
$6 \cdot 10^{-6}/16$ was chosen to obtain the predicted capillary
number $0.05$. The forcing for other grids can be obtained using
simple proportionality relationships:
\begin{equation}
\begin{aligned}
&\Ca_{\mathrm{lit}} \propto U_{\mathrm{bubble}}\\
&U_{\mathrm{bubble}} \propto \frac{\mathrm{d}P}{\mathrm{d}x} N_y^2\\
&\Ca_{\mathrm{lit}} \propto \frac{\mathrm{d}P}{\mathrm{d} x} N_y^2 \text{ or }\\
&\frac{\mathrm{d}P}{\mathrm{d} x} \propto \frac{\Ca_{\mathrm{lit}}}{N_y^2},
\end{aligned}
\end{equation}
where the subscript ,,lit'' stands for the predicted capillary number
\cite{giavedoni-numerical,heil-bretherton}. The grid number $N_y$ is not
involved because all simulations are conducted on the same grid.
The pressure gradient can be obtained through the capillary number
ratio:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d} x}=6 \cdot 10^{-6}/16 \frac{\Ca_{\mathrm{lit}}}{0.05}
\end{equation}
The film thickness is initialized through the ratio of capillary numbers as well:
\begin{equation*}
w=12 \frac{\Ca_{\mathrm{lit}}}{0.05}
\end{equation*}
The results obtained after $2\cdot10^5$ steps are presented in Table
\ref{table:parameters:capillary:number}. One can see that the calculated
capillary numbers are overpredicted (the actual capillary number is larger than the capillar number
we aim for, $\Ca>\Ca_{\mathrm{lit}}$). We
extracted
the data of 
\citet{giavedoni-numerical} and \citet{heil-bretherton} with the help of software ``Engauge
Digitizer'' and compared them with our results (see 
Fig. \ref{fig:capillary:comparison}).
\begin{figure}
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison.eps}
\caption{A comparison between simulation results and results of
\citet{giavedoni-numerical} and \citet{heil-bretherton}. One can see a
reasonable agreement. The plots depict the film thickness as a function of the
the capillary number.\label{fig:capillary:comparison}}
\end{figure}

@@

\subsection{3D results}
One can see from Fig. \ref{fig:heil:three:dim} that the difference between the axial and diagonal
radii
happens at the capillary numbers less than $0.1$. However the axial radius has a value around $0.99
H_{\mathrm{eff}}$. Therefore, the film occupies $0.01 H_{\mathrm{eff}}$ which needs to be resolved
with at minimum $10$ lattice units (see results for grid independency). This condition implies grids
to be at least size $1000\times1000\times 15000$, which is not a computationally feasible
task. Even the symmetric and non-symmetric profiles are obtained by current simulations, Fig.
\ref{fig:sym:antisym}, the grid independence is only guaranteed for symmetric cases.
Therefore our target simulations are for $\Ca\geq 0.1$, where the axial radius equals to the
diagonal radius. In what follows we will show two particular simulations. One is for the dependance
of the bubble radius on the capillary number. Another one is for the velocity patterns. In
comparison with simulations of \citet{heil-threedim} the transition capillary
number is located in the range $0.47<\Ca<0.63$. 
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/sym.eps}\\
\includegraphics[width=0.5\textwidth]{Figures/antisym.eps}\\
\caption{Symmetric ($\Ca=0.63$) and non-symmetric ($\Ca=0.026$) crosssections in the middle of the
bubble. One can see that for the symmetric profile $R_h=R_d$ and for the antisymmetric $R_h\neq
R_d$.  \label{fig:sym:antisym} }
\end{figure}

We also simulate only the quarter of the channel by imposing the mirror boundary conditions. This
gives an improvement in memory requirements and simulution time by factor of four.  
\subsubsection{Capillary region}
We performed a number of simulations of grid size $52\times52\times 1500$ for the
quarter of geometry and of refined grid size $82\times82\times 1500$. Other parameters were
chosen as
$k=0.04$,$A=0.04$, and $\Gamma=1.0$.
